---
title: "Job Market Hub"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(lubridate)
library(googlesheets4)
library(kableExtra)
library(DT)
library(fontawesome)

## package global options
# skip googlesheets login
gs4_deauth()

```

## Welcome!

This page is for internal use only. If you are not a letter writer, I encourage you to use the navigation tab on top to learn more about me and my work.

## Upcoming Deadlines

```{r, layout="l-body-outset"}
jobs = read_sheet("https://docs.google.com/spreadsheets/d/1wCTQQpg9AALD-d4N2gCD-_xrV_P-EhNJUeDQnUsVHPI/edit#gid=505775282")

# some manipulation to show only what is useful
jobs = jobs %>% 
  filter(is.na(applied)) %>% 
# supressing filter by date for now
#  filter(as_date(deadline) < as_date(Sys.Date() + ddays(30))) %>% 
  mutate(
  # url = paste0('<a href="', url,'"target="_blank"></a>'),
  deadline = format(deadline, "%B %d, %Y")
)


## Saving these lines in case the filterable table returns
# select(deadline, field, place, department, position, letters, Jake, Matt, Michelle, Gisela, Avital, notes, url) 
  
  
  # datatable(options = list(scrollX = TRUE), 
  #           escape = FALSE, rownames = FALSE, 
  #           class = 'hover stripe nowrap',
  #           filter = list(
  #             position = 'top', clear = FALSE))


```

**Last update:** `r format(Sys.Date(), "%B %d, %Y")`. Full list [here](https://docs.google.com/spreadsheets/d/1wCTQQpg9AALD-d4N2gCD-_xrV_P-EhNJUeDQnUsVHPI/edit#gid=505775282).

**Number of letters to write:**

```{r}
jobs %>% 
  summarize(across(Avital:Michelle, ~ sum(.x, na.rm = TRUE))) %>% 
  select_if(~sum(.) > 0) %>% 
  kbl()
```

```{r, results = "asis"}
refs = jobs %>% select(Avital:Michelle)

# try a loop now
for(i in 1:nrow(jobs)){
  cat("[**",
      jobs$position[i], " - ",
      jobs$department[i], " - ",
      jobs$place[i],
      paste0('**](', jobs$url[i],')'),
      "<br/>",
      sep = "")
  
  cat("**Letters:**",
      paste(colnames(refs)[which(refs[i,] == 1)], collapse = ", "),
      "<br/>",
      "**Deadline:**", jobs$deadline[i],
      "<br/>",
      "**Delivery:**", jobs$letters[i],
      "<br/>",
      "**Notes:**", jobs$notes[i],
      "<br/>")
  
  cat(rep("&nbsp;  ", 2), sep="\n")
}
```


## Materials

- Cover letter [`[Comparative]`](job_market/cover_letter/cover_letter_cp.pdf) [`[Methods]`](job_market/cover_letter/cover_letter_methods.pdf)

- Research statement [`[Comparative]`](job_market/statements/research_statement_cp.pdf) [`[Methods]`](job_market/statements/research_statement_methods.pdf)

- Teaching statement [`[Comparative]`](job_market/statements/teaching_statement_cp.pdf) [`[Methods]`](job_market/statements/teaching_statement_methods.pdf)

## Writing Samples

Coming soon!

## Teaching

- [Syllabi](/teaching.html)

- [Teaching evaluations](/job_market/statements/diaz_teaching_evals.pdf)
