---
title: "Job Market Hub"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(lubridate)
library(googlesheets4)
library(DT)
library(fontawesome)

## package global options
# skip googlesheets login
gs4_deauth()

```

## Welcome!

You are here either because you are a recommendation letter writer or because someone pointed you out to my job market materials. This page is semi-public and mostly for internal use, but feel free to share if it's helpful.

## Upcoming Deadlines

**Last update:** `r format(Sys.Date(), "%B %d, %Y")`. Full list [here](https://docs.google.com/spreadsheets/d/1wCTQQpg9AALD-d4N2gCD-_xrV_P-EhNJUeDQnUsVHPI/edit#gid=505775282).

```{r, layout="l-body-outset"}
jobs = read_sheet("https://docs.google.com/spreadsheets/d/1wCTQQpg9AALD-d4N2gCD-_xrV_P-EhNJUeDQnUsVHPI/edit#gid=505775282")

# some manipulation to show only what is useful
jobs %>% 
  filter(is.na(applied)) %>% 
# supressing filter by date for now
#  filter(as_date(deadline) < as_date(Sys.Date() + ddays(30))) %>% 
  mutate(
  Jake = ifelse(Jake == 1, "Yes", "No"),
  Matt = ifelse(Matt == 1, "Yes", "No"),
  Michelle = ifelse(Michelle == 1, "Yes", "No"),
  Gisela = ifelse(Gisela == 1, "Yes", "No"),
  Avital = ifelse(Avital == 1, "Yes", "No"),
  url = paste0('<a href="', url,'"target="_blank">Link</a>'),
  deadline = format(deadline, "%m-%d-%Y")
) %>% 
  select(deadline, field, place, department, position, letters, Jake, Matt, Gisela, Avital, notes, url) %>% 
  datatable(options = list(scrollX = TRUE), 
            escape = FALSE, rownames = FALSE, 
            class = 'hover stripe nowrap',
            filter = list(
              position = 'top', clear = FALSE))


```


## Materials

- Cover letter [`[Comparative]`](job_market/cover_letter/cover_letter_cp.pdf) [`[Methods]`](job_market/cover_letter/cover_letter_methods.pdf)

- Research statement [`[Comparative]`](job_market/statements/research_statement_cp.pdf) [`[Methods]`](job_market/statements/research_statement_methods.pdf)

- Teaching statement [`[Comparative]`](job_market/statements/teaching_statement_cp.pdf) [`[Methods]`](job_market/statements/teaching_statement_methods.pdf)

## Writing Samples

Coming soon!

## Teaching

- [Syllabi](/teaching.html)

- [Teaching evaluations](/job_market/statements/diaz_teaching_evals.pdf)
