---
title: "Cumulative Learning"
subtitle: "Politics of Developing Democracies"  
date: "April 26-28, 2022"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%/%total%" 
      highlightStyle: github
      highlightLines: true
      ratio: "16:9"
      countIncrementalSlides: false
---

```{r setup, include = FALSE, warning = FALSE}
# Loads knitr and xaringan themer settings
source("theme.R")
```

```{r other-options}
library(tidyverse)
library(kableExtra)
library(fontawesome)

# Only for this week
library(DeclareDesign)

# ggplot global options
theme_set(theme_bw(base_size = 20))
```

class: inverse

## End of semester reminders

- Final paper due on **Saturday, May 7, 12:00 PM**

- Your paper will be graded by whoever was in charge of the class when you presented

- Available for meetings/feedback during May 2-6 `(email before scheduling)` 

- Final grades due on **Tuesday, May 17** `(May 15 if graduating)`

- Some flexibility at the margins, but stay accountable

---
class: inverse

## Roadmap for this week

- **Today:** Take a step back, think about what we can learn from policy interventions that emphasize research designs for causal inference

- **Thursday:** Open-ended conversation about coordinated trials and what you learned in this course `(read one of the required readings)` [`r fa("external-link-alt")`](#thursday)

---

## Learning from experiments

- How do you prove that a policy interventions works?

--

- We want to make statements about causation `(causal inference)`

    - *CCTs improve health and education*
    - *Monitoring reduces corruption*
    
--

- To back-up those statements, we need to rule out potential **confounding factors**

    - *Healthier, educated families are more likely to meet the CCT requirements*
    - *Monitoring is only possible in less corrupt places*
    
--

- One way to rule out potential confounders is to conduct an experiment or analyze existing data that looks like an experiment `(RDD, natural experiment, quasi-experiment)`

- **Challenge:** This is only true *in expectation*

---
## A small experiment

```{r}
pop = declare_population(
  N = 4,
  female = c(0, 0, 1, 1),
  Y0 = c(0, 0, 0, 1),
  Y1 = c(0, 1, 1, 1)
)

pot = declare_potential_outcomes(Y ~ Y1 * Z + Y0 * (1-Z))

estimand = declare_inquiry(ATE = mean(Y1 - Y0))

assign = declare_assignment(Z = complete_ra(N, m = 2))

reveal = declare_measurement(Y = reveal_outcomes(Y ~ Z))

estimator = declare_estimator(Y ~ Z, inquiry = "ATE")

design_1 = pop + pot + estimand + assign + reveal + estimator
```


```{r}
set.seed(141)

dat = draw_data(design_1)

dat_0 = dat %>% select(ID, female, Y0, Y1)

colnames(dat_0) = c("ID", "Female", "Y(0)", "Y(1)")

dat_0 %>% kbl()
```

- $Y(*)$ are the **potential outcomes** under control `(0)`  and treatment `(1)`, respectively

- $Y(*) = 1$ means person's life improves, $Y(*) = 0$ means life stays the same

--

- We have:

    - One person for which the policy would do nothing
    - Two people for which the policy improves life
    - One person who improves their life either way
    
---
## Assign policy treatment at random

```{r}
dat_1 = dat %>% select(ID, female, Y0, Y1, Z)

colnames(dat_1) = c("ID", "Female", "Y(0)", "Y(1)", "Z")

dat_1 %>% kbl()
```

--

- We happened to randomly assign the policy to the two women `(I did not mean this to happen in the first trial)`

- We only observe the potential outcomes that corresponds to the treatment status

---
## Revealing outcomes

```{r}
dat_2 = dat %>% select(ID, female, Y0, Y1, Z, Y)

colnames(dat_2) = c("ID", "Female", "Y(0)", "Y(1)", "Z", "Y obs")

dat_2 %>% kbl()
```

--

- The **true** treatment effect is 

$$ATE = E[Y(1)] - E[Y(0)] = 3/4 - 1/4 = 1/2$$

- Which we **cannot observe in the real world**

--

- We can **approximate** the ATE with $\widehat{ATE} = 2/2 - 0/2 = 1$

- We are off the mark! What happens if we redo the experiment?

---

## Redoing the experiment

```{r}
set.seed(142)

redo = draw_data(design_1)

dat_3 = redo %>% select(ID, female, Y0, Y1, Z, Y)

colnames(dat_3) = c("ID", "Female", "Y(0)", "Y(1)", "Z", "Y obs")

dat_3 %>% kbl()
```

- New treatment assignments but the same revealed outcomes `(this was an accident)`

- We still have $ATE = 1/2$

- But now $\widehat{ATE} = 1/2 - 1/2 = 0$

- Which is off the mark in the opposite direction

---
## Why does this happen?

```{r}
dat_4 = dat %>% select(ID, female, Y0, Y1, Z, Y)

dat_4$Z2 = dat_3[, 5]

dat_4$Y2 = dat_3[, 6]


colnames(dat_4) = c("ID", "Female", "Y(0)", "Y(1)", "Z", "Y obs", "Z", "Y obs")

dat_4 %>% kbl() %>% add_header_above(c(" " = 4, "Experiment 1" = 2, "Experiment 2" = 2))
```

- Perhaps men and women, as groups, react to the policy differently

- We want to rule out that the results of our experiment depend on whether we assign treatments to men or women

- **Experiment 1:** 2/2 women in treatment and 0/2 in control `(imbalanced)`

- **Experiment 2:** 1/2 woman in treatment and 1/2 in control `(balanced)`

--

- Does that mean that experiment 2 is free from **random confounding**?
--
 Maybe
 
---
## Redo 1,000 experiments

.center[
```{r, out.width = "60%", cache = TRUE}
set.seed(20220425)
sims = simulate_design(design_1, sims = 1000)

ggplot(sims) +
  aes(x = estimate) +
  geom_vline(xintercept = mean(sims$estimand), linetype = "dashed", size = 2) +
  geom_density(size = 2) +
  labs(x = "Observed ATEs", y = "Density") +
  annotate("text", x = mean(sims$estimand) + 0.1, y = 1.5, label = "True ATE", 
           size = 6, color = "#336666") +
  theme_xaringan()
```
]

--

- About half of the time we get the true ATE spot on

- The other half we are off in either direction

---
## What does this mean?

- Experiments only rule out the role of potential confounders **IN EXPECTATION**

- We can sustain this claim in two ways

--

1. With a sufficiently large sample `(But how large is large enough?)`
    
2. By repeating the same experiment multiple times `(Nobody does this and we can't even if we tried)`

--

- Our confidence in the results of an experiment depends on two dimensions of its research design

1. **Bias:** In average, how far off we are from the truth

2. **Precision:** How often do we get the same result

---

## Visualizing bias and precision

.center[
```{r, out.width = "90%"}
d1 = data.frame(
  x = rnorm(100, 0, 1),
  true = 0,
  val = "Unbiased",
  rel = "Precise"
)

d2 = data.frame(
  x = rnorm(100, 0, 5),
  true = 0,
  val = "Unbiased",
  rel = "Imprecise"
)

d3 = data.frame(
  x = rnorm(100, 5, 1),
  true = 0,
  val = "Biased",
  rel = "Precise"
)

d4 = data.frame(
  x = rnorm(100, 5, 5),
  true = 0,
  val = "Biased",
  rel = "Imprecise"
)

d = rbind(d1, d2, d3, d4)

d$val = fct_relevel(d$val, "Biased", "Unbiased")

d$rel = fct_relevel(d$rel, "Precise", "Imprecise")

ggplot(d) +
  aes(x = x) +
  geom_density(size = 1) +
  geom_vline(xintercept = 0, size = 2, color = "#336666", linetype = "dashed") +
  annotate("text", x = - 5, y = 0.38, label = "True ATE", color = "#336666", size = 5) +
  labs(x = "Observed ATEs", y = "") +
  facet_grid(rel ~ val) +
  theme_xaringan() +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank())
```
]

---
## In practice

- We only have a decent guess of where we are in terms of bias and precision, but we do not know for sure

- So you need a lot of domain expertise to justify that the ATE you observed can be attributed to the policy

- This often involves showing not only whether something works but also **explaining why it works**

- This is the first step towards knowing whether it would **work somewhere else**

---
## Why do interventions work?

- If everything goes perfectly with the research design, we can only determine whether **what works**

- To turn the results of an experiment into useful knowledge, we need to know **why** it works

--

- Understanding why something works requires domain expertise

--

- However, experiments assume as little as possible about a study. All we need to know is that treatments randomly assigned

--

- Experimentation and domain expertise tend to be at odds, and this is a problem we have to deliberately address!

---
## Generalization and extrapolation

- One critique to experiments is that they invest in **internal validity** at the expense of **external validity**

--

- **Internal validity:** We can `(probabilistically)` attribute the observed effect to the policy intervention

- **External validity:** Whether the effect **extrapolates** or **generalizes**

- **Extrapolation:** Whether it works *elsewhere* `(in different cases)`

- **Generalization:** Whether it works *everywhere* `(in set of cases)`

--

**Deaton and Cartwright (2018):** This is reductive, we tend to think that interventions travel or not, but there is more to it

---
## Support factors

- **Example:** A house burns down because the television was left on

--

- Not all houses with TVs left on burn down, but sometimes they do, perhaps because the wiring was poor

--

- A **support factor** is one part of the **causal pie**

--

- **Causal pie:** A set of causes that are jointly but not separately sufficient for a contribution to an effect `(Social scientists call this INUS causation)`

--

- **Analogy:** CCT transfers only improve health if we have decent clinics

---

## Scales and drills

-  Even if we understand why something works, we can still have questions about the applicability of a policy

--

- **Scaling up:** Whether we can apply the intervention to a much broader area 

    - Small scale interventions can become **unfeasible** or **cost-prohibitive** in a larger scale

    - Even if viable, some policies only work at a small scale!

--

- **Drilling down:** Can we apply the results of an intervention to individual units?

    - Just because an intervention works **in average**, it does not mean that everyone will benefit from them

    - **Example:** Not every woman will vote freely when encouraged to do so

    - Most of the time, using experiments to evaluate policies means wasting money on people for whom the policy does not work

    - This can be **unethical** if it means preventing people from accessing the help they need
    
---
class: inverse

## Takeaways

- Never takes experiments at face value

- They are useful tools to evaluate policies, but they do not automatically teach us what works

- We still have to answer questions about:

    - **Statistical inference:** Does the research design yield unbiased and precise estimates?
    
    - **Causal inference:** Do we understand what works and why it works?
    
    - **Generalization** and **extrapolation**
    
- **Thursday:** One solution to overcome all challenges at the same time is using **coordinated trials**
