---
title: "Economic Determinants of Democracy and Dictatorship"
subtitle: "Comparative Politics"  
date: "October 11-15, 2021"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%/%total%" 
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)

library(knitr)
opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
# In the future you want to move this to a separate file and source it every time you create a new file
library(xaringanthemer)
style_duo_accent(
  title_slide_background_image = "figs/logo.png",
  title_slide_background_size = "8%",
  title_slide_background_position = "50% 95%",
  primary_color = "#336666",
  secondary_color = "#71C5E8",
  inverse_header_color = "#FFFFFF",
  background_color = "#EAE9EA",
  link_color = "#71C5E8",
  # easy to fetch colors
  colors = c( 
    white = "#FFFFFF",
    green = "#336666",
    lblue = "#71C5E8"
    )
)
```

```{r other-options}
library(tidyverse)
library(kableExtra)
library(fontawesome)

# ggplot global options
theme_set(theme_bw(base_size = 20))

# Plot democracy and WDI data
library(countrycode)
library(WDI)
# remotes::install_github('rensa/ggflags')
library(ggflags)
```

## Last time

- We discussed classifying regime types

    - Continuous or categorical
    - Minimalist or substantive

- This all matters because we want to understand the consequences of being a democracy or a dictatorship

- We usually promote democracy because we believe it brings good outcomes

- But how do countries become democracies?

- And once they do, how do they stay democracies?

- This week is about the **economic** explanations for regime change and survival

- Next week is about the **cultural** explanations

---

## Modernization theory

- All societies pass through the same stages of economic development

.pull-left[
### Traditional society

- Large agriculture
- Small industry
- Small service
- Dictatorship
]

.pull-right[
### Modern society

- Small agriculture
- Large industry
- Large service
- Democracy
]

- As societies develop, they are `more likely` to become a democraty
- So we say that *development* causes *democracy*

---
class: inverse 

## Causal statements in the social sciences

.center[
### Economic development $\rightarrow$ Democracy
]

--

.center[
### Explanatory variable $\rightarrow$ Outcome variable
]

--

.center[
### X $\rightarrow$ Y
]

--

- We do not see `causation` happening, only `association`

--

- So just because richer countries tend to be democracies it does not mean that development `causes` democracy

--

- It could be the other way around

--

- Or a big coincidence!

--

- Hence the importance of finding **critical tests** for our theories

---

## Using DD index

```{r, fig.height=5}
load("data/dem_scores.RData")

gdp = WDI(indicator = "NY.GDP.PCAP.KD", country = "all", start = 2007, end = 2020)

# some recoding
gdp$iso3c = countrycode(gdp$iso2c, origin = "iso2c", destination = "iso3c")

gdp$gdp_pc = gdp$NY.GDP.PCAP.KD/1000

# merge
dd_gdp = left_join(dd, gdp %>% filter(year == 2007), by = "iso3c")

# visualize
dd_gdp_df = dd_gdp %>% 
  group_by(dd) %>% 
  summarize(gdp_avg = mean(gdp_pc, na.rm = TRUE))


ggplot(dd_gdp_df) +
  aes(x = as.factor(dd), y = gdp_avg) +
  geom_col() +
  labs(x = "2008 DD Index", y = "mean GDP per capita in\n2000 PPP USD (thousands)") +
  scale_x_discrete(labels = c("Dictatorships", "Democracies")) +
  theme_xaringan()
```

---

## Using Polity

```{r, fig.height = 5}
gdp$iso2c = tolower(gdp$iso2c)

polity_gdp = left_join(polity, gdp %>% filter(year == 2017), by = "iso3c")

# Kosovo is messing things up
polity_gdp = polity_gdp %>% 
  filter(polity_annual_country != "Kosovo") %>% 
  mutate(polity = ifelse(abs(polity) > 10, NA, polity))

ggplot(polity_gdp) +
  aes(x = log(gdp_pc), y = polity) +
  geom_flag(aes(x = log(gdp_pc), y = polity, country = iso2c)) +
  geom_smooth(size = 2, se = FALSE) +
  labs(x = "GDP per capita (logged)", y = "Polity score") +
  theme_xaringan()
```

---

## Using Freedom House

```{r, fig.height = 5}
fh_gdp = left_join(fh, gdp %>% filter(year == 2019), by = "iso3c")

# drop duplicate countries
fh_gdp = fh_gdp %>%  filter(fh_country != "Micronesia" & fh_country != "Kosovo")

ggplot(fh_gdp) +
  aes(x = log(gdp_pc), y = fh_total_reversed) +
  geom_flag(aes(x = log(gdp_pc), y = fh_total_reversed, country = iso2c)) +
  geom_smooth(size = 2, se = FALSE) +
  labs(x = "GDP per capita (logged)", y = "FH score (reversed)") +
  theme_xaringan()
```

---

## Using V-DEM

```{r, fig.height = 5}
vdem_gdp = left_join(vdem, gdp %>% filter(year == 2019), by = "iso3c")

# drop duplicates
vdem_gdp = vdem_gdp %>%
  filter(country_name != "Zanzibar" & country_name != "Kosovo") %>% 
  filter(!is.na(iso3c))

ggplot(vdem_gdp) +
  aes(x = log(gdp_pc), y = v2x_polyarchy) +
  geom_flag(aes(x = log(gdp_pc), y = v2x_polyarchy, country = iso2c)) +
  geom_smooth(size = 2, se = FALSE) +
  labs(x = "GDP per capita (logged)", y = "V-DEM polyarchy score") +
  theme_xaringan()
```

---

## Two interpretations

- The positive association between economic development and democracy is consistent with **two explanations**

1. **Modernization Theory (Lipset):**

    - Development makes democracy `more likely to emerge` **AND**
    - Development makes democracy `more likely to survive`
    
2. **Survival Story (Przeworski):**

    - Development makes democracy `neither more nor less likely to emerge` **BUT**
    - Development makes democracy `more likely to survive`
 
--
    
- **CGG** show support for modernization theory by looking at transition probabilities

- Let's look at it from a different angle

---
## Revisiting the survival story

```{r, fig.height = 5}
polity_gdp = polity_gdp %>% 
  mutate(group = ifelse(polity > 0, 1, 0))

ggplot(polity_gdp) +
  aes(x = log(gdp_pc), y = polity, color = as.factor(group)) +
  geom_hline(yintercept = 0, size = 1, linetype = "dashed") +
  geom_flag(aes(x = log(gdp_pc), y = polity, country = iso2c)) +
  geom_smooth(size = 2, se = FALSE) +
  labs(x = "GDP per capita (logged)", y = "Polity score") +
  theme(legend.position = "none") +
  scale_color_manual(values = c("#71C5E8", "#336666")) +
  theme_xaringan()
```

---

## Revisiting the survival story

- `Rich democracies` are "more democratic"
- `Rich dictatorships` are "more autocratic"

- Which supports the survival story! More money makes any regime more stable

&nbsp;

--

- This is not to say that **CGG** are wrong

- Rather, that this is an open debate and the devil is on the details

- How we `define` and `operationalize` development and democracy matters a lot

--
```
(Research methods matter a lot too, but this course is not about that...)
```

---

## Critiques to development $\rightarrow$ democracy

[CONTINUE HERE DEPENDENCY THEORY]