---
title: "Democracy and Dictatorship"
subtitle: "Comparative Politics"  
date: "October 4-8, 2021"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%/%total%" 
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)

library(knitr)
opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  title_slide_background_image = "figs/logo.png",
  title_slide_background_size = "8%",
  title_slide_background_position = "50% 95%",
  primary_color = "#336666",
  secondary_color = "#71C5E8",
  inverse_header_color = "#FFFFFF",
  background_color = "#EAE9EA",
  link_color = "#71C5E8",
  # easy to fetch colors
  colors = c( 
    white = "#FFFFFF",
    green = "#336666",
    lblue = "#71C5E8"
    )
)
```

```{r other-options}
library(tidyverse)
library(kableExtra)
library(fontawesome)

# ggplot global options
theme_set(theme_bw(base_size = 20))

# Democracy scores maps
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(countrycode)
library(gridExtra)
```

class: inverse

## Reminder

- Quiz 1 opens on Thursday, October 7 at noon
- Due on Friday, October 8 at 5:00 PM
- Weeks 1-4

---

## Last time

- We study politics within countries, by which we usually mean nation-states
- Two explanations for the emergence of modern nation-states:

    1. Contractarian view
    2. Predatory view
    
- They also help us understand variation in **state capacity**, **state fragility**, or **stateness** across countries and over time

- This is the foundation for many other interesting questions

---

## The way forward

- Why are some countries `democracies` and other `dictatorships`?

- Which `regime` produces better economic performance?

- What factors influence democratic survival?

&nbsp;

- To answer these questions, we need to classify countries as **democracies** or **dictatorships**

---
class: inverse

## Aside: English is hard

- **State:** An organization with comparative advantage in violence

- **Government:** The group of people who rule the state

- **Regime:** Institutions that determine how governments are formed

--

&nbsp;

.center[This week is about **regime types**]

---
layout:true 
## Democracy is an old concept

---
- ... and it was a bad thing!

- **Demokratia:** Rule by the `(common)` people

- **Plato:** Government by the poor, uneducated against the rich, educated

- Would lead to mob rule and class warfare

---

- **Aristotle:** All forms of government are corruptible

```{r}
aristotle = data.frame(
  number = c("One", "Few", "Many"),
  good = c("Monarchy", "Aristocracy", "Politeia"),
  bad = c("Tyranny", "Olygarchy", "Democracy")
)

colnames(aristotle) = c("Number of rulers", "Good for all", "Good for the rulers")

aristotle %>% 
  kbl() %>% 
  add_header_above(c(" " = 1, "Good form" = 1, "Bad form" = 1))
```

- **Democracy:** Rule by the worst class (Monarchy preferred)
- **Until 18th Century:** 

    - No elections `(aristocracy)`, but `lottery`
    - No representation, but `direct legislation`
    
- `Democracy` $=$ `representative government` only after French and American Revolutions

---

- As demands for representative government increased, the classical 3-way distinction disappeared in favor of a 2-way distinction between **democracy** and **dictatorship**

- This is also when democracy started to associate with `equality`, `justice`, and `freedom`

---
layout:false

## Defining democracy

- These are **substantive** definitions of democracy (and dictatorship) since they focus on what it `should accomplish` rather than what `it is`

--

- **Robert Dahl** proposed a **minimalist** definition instead

- He used the term **polyarchy** to distinguish it from conventional use. Also to emphasize that this as an ideal type, and that different regimes can get closer or further from the ideal

---

## Polyarchy

- Regimes can be categorized in two dimensions:

    1. **Contestation:** Extent to which citizens can organize into competing blocks to press for policy change
    2. **Inclusion:** Who gets to participate

```{r}
polyarchy = data.frame(
  countries = c("Apartheid\nSouth Africa", "China", "Soviet Union", "United States", "Polyarchy\n(Ideal type)"),
  Inclusion = c(10, 30, 60, 70, 90),
  Contestation = c(70, 10, 10, 70, 90)
)

ggplot(polyarchy) +
  aes(x = Inclusion, y = Contestation, label = countries) +
  geom_text(size = 5) +
  scale_x_continuous(breaks = c(10, 90), labels = c("Low", "High")) +
  scale_y_continuous(breaks = c(10, 90), labels = c("Low", "High")) +
  coord_cartesian(xlim = c(-5, 105), ylim = c(-5, 105)) +
  theme_xaringan() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
 

```


---

## Polyarchy

- This definition is **minimalist** because its sole purpose is to classify regimes

- We want this because later on we want to learn about the causes and consequences of regime types

- `Contestation` and `inclusion` are abstract concepts

- How do we turn them into a concrete measure of democracy?

---

## From concept to measurement

- **Concept:** Mental category that conveys meaning. Like `theories` or `models` they are always wrong, but sometimes useful

--

- **Measure/indicator:** An observable representation of a concept

    - CGG say a `quantification`, but a measure need not be a "quantitiy" `r fa("fire-alt")`
    - `Example:` Hot and cold are categorial measures of temperature
    
--

- **Operationalization:** The process by which we translate a concept into a measure. It often implies justifying why the measure is a good representation of the concept

---

## Measures of Democracy

.pull-left[

### Substantive
- Freedom House [`r fa("external-link-alt")`](https://freedomhouse.org/)

- Varieties of Democracy [`r fa("external-link-alt")`](https://www.v-dem.net/en/)
]

.pull-right[

### Minimalist
- Polity IV-V [`r fa("external-link-alt")`](http://www.systemicpeace.org/polityproject.html)

- Democracy-Dictatorship Index [`r fa("external-link-alt")`](https://sites.google.com/site/joseantoniocheibub/datasets/democracy-and-dictatorship-revisited)
]

--

&nbsp;

.center[Let's talk about these in logical order]

---
## Freedom House

- Based on two dimensions:

    1. Political rights (10 questions, scored 0-4)
    2. Civil rights (15 questions, scored 0-4)

- Both transformed to a 7 point scale
- A country's FH is the average of political and civil rights
- A `substantive` measure because democracy should promote freedom

---

## Polity IV-V

- $Polity = Democracy - Autocracy$

- Ranges from -10 to 10

- Five dimensions:

    1. Competitiveness of executive recruitment
    2. Openness of executive recruitment
    3. Regulation of political participation
    4. Competitiveness of political participation
    5. Executive constraints

- `Minimalist` since 1 to 4 capture `inclusion` and `contestation`
- `Executive constraints` adds the idea that governments must be limited. Most of the variation in Polity scores depends on it


---

## Democracy-Dictatorship Index (DD)

- Doubles down on the `minimalist` approach

- Claims the core feature of democracy is `contestation`

- A country is a `democracy` if:

    1. The chief executive is elected
    2. The legislature is elected
    3. There is more than one party competing in elections
    4. An alternation in power under identical electoral rules has taken place
    
- If any of these conditions does not hold, the country is a `dictatorship`

- This is a **dichotomous** or **binary** measure



---

## Varieties of Democracy (V-DEM)

- Accepts that `different definitions` of democracy have `different purposes`
- Takes a `modular` approach. You can build your own definition

- Country experts answer questions in `five dimensions`

    1. **Electoral democracy** (polyarchy)
    2. **Liberal democracy** (individual freedoms, checks and balances)
    3. **Participatory democracy** (civil society organizations, direct democracy, local government)
    4. **Deliberative democracy** (engaged society)
    5. **Egalitarian democracy** (equal access, protection, distribution)
    
- Each one of these is a score ranging from 0 to 1 (1 being the ideal type)
- **Transparency:** We can combine or disaggregate scores depending on our goals

---
## What's a good measure?

- The perfect measure is **reliable** and **valid**

- **Validity:** Does it capture the concept we want to measure?

    - Requires us to imagine there is a "true" measure that we cannot observe
    - How close can we get with different approaches?
 
    
--

- **Reliability:** Is the measurement consistent?

    - Requires us to imagine the possibility of repeated measures `(e.g. multiple experts or coders)`
    - Do we get the same measurement over repeated measures?
    
--

- We never observe **validity** but we can make hypothetical statements about it
- We can observe **reliability** for scores reporting the answers of individual experts (V-DEM)

---
## Visualizing validity and reliability

```{r, fig.height=5}
d1 = data.frame(
  x = rnorm(100, 0, 1),
  true = 0,
  val = "Valid",
  rel = "Reliable"
)

d2 = data.frame(
  x = rnorm(100, 0, 5),
  true = 0,
  val = "Valid",
  rel = "Unreliable"
)

d3 = data.frame(
  x = rnorm(100, 5, 1),
  true = 0,
  val = "Invalid",
  rel = "Reliable"
)

d4 = data.frame(
  x = rnorm(100, 5, 5),
  true = 0,
  val = "Invalid",
  rel = "Unreliable"
)

d = rbind(d1, d2, d3, d4)

d$val = fct_relevel(d$val, "Valid", "Invalid")

ggplot(d) +
  aes(x = x) +
  geom_density(size = 1) +
  geom_vline(xintercept = 0, size = 2, color = "#336666", linetype = "dashed") +
  annotate("text", x = - 3, y = 0.38, label = "Truth", color = "#336666", size = 5) +
  labs(x = "Repeated measures", y = "") +
  facet_grid(rel ~ val) +
  theme_xaringan() +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank())
```


---
layout: true

## Does it matter?

---

```{r}
load("data/dem_scores.RData")

# fix polity score
polity = polity %>% mutate(polity = ifelse(abs(polity) > 10, NA, polity))

# Merge with map
world = ne_countries(scale = "small", returnclass = "sf")

fh_map = left_join(world, fh, by = c("iso_a3" = "iso3c"))
poli_map = left_join(world, polity, by = c("iso_a3" = "iso3c"))
dd_map = left_join(world, dd, by = c("iso_a3" = "iso3c"))
vdem_map = left_join(world, vdem, by = c("iso_a3" = "iso3c"))

fhp = ggplot(fh_map) +
  geom_sf(aes(fill = fh_total_reversed)) +
  labs(fill = "") +
  theme_void() +
  theme(legend.position = "bottom") +
  scale_fill_viridis_c()

polip = ggplot(poli_map) +
  geom_sf(aes(fill = polity)) +
  theme_xaringan() +
  labs(fill = "") +
  theme_void() +
  theme(legend.position = "bottom") + 
  scale_fill_viridis_c()

ddp = ggplot(dd_map) +
  geom_sf(aes(fill = as.factor(dd))) +
  theme_xaringan() +
  labs(fill = "") +
  theme_void() +
  theme(legend.position = "bottom") +
  scale_fill_viridis_d(labels = c("Dictatorship", "Democracy", "NA"))

vdemp = ggplot(vdem_map) +
  geom_sf(aes(fill = v2x_polyarchy)) +
  theme_xaringan() +
  labs(fill = "") +
  theme_void() +
  theme(legend.position = "bottom") +
  scale_fill_viridis_c()
```

### Freddom House 2020 (reversed)

```{r, fig.dim = c(6,3)}
fhp
```

---

### Polity V 2018

```{r, fig.dim = c(6,3)}
polip
```

---

### DD Index 2008

```{r, fig.dim = c(6,3)}
ddp
```

---

### V-DEM Polyarchy score 2020

```{r, fig.dim = c(6,3)}
vdemp
```

---

### Comparing scores

- Do these maps look similar or different to you?

- FH, Polity, V-DEM rely on expert judgments

- DD appears as a more straightforward coding scheme but it gets tricky as the margins `(e.g. Is the election really competitive?)`

- V-DEM highly transparent, but people can still use it wrong

- Most scholarship uses one and then reports similar results with other scores

- So, does the substantive vs. minimalist distinction make sense if people use these interchangeably? `r fa("fire-alt")`

---
layout: false

## Takeaways

- Democracy used to mean `(bad) government by the poor, uneducated`

- It only became to mean representative government, equality, justice, and freedom after the French and American revolutions

- At the same time, the democracy vs. dictatorship dichotomy appears

- Regime type matters to understand a lot of political, social, and economic outcomes. So we care about measurement

- We can measure regimes with substantive or minimalist definitions

- Substantive definitions are good to identify whether regimes are delivering what they promise

- Minimalist definitions are good to understand the causes and consequences of regime types

- We want our measures to be `valid` and `reliable`


---
class: inverse center middle

# Reminder:
## Quiz 1 Due Friday by 5:00 PM

# Next Week:
## Economic Determinants of Democracy and Dictatorship