---
title: "Democracy and Dictatorship"
subtitle: "Comparative Politics"  
date: "October 4-8, 2021"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%/%total%" 
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)

library(knitr)
opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  title_slide_background_image = "figs/logo.png",
  title_slide_background_size = "8%",
  title_slide_background_position = "50% 95%",
  primary_color = "#336666",
  secondary_color = "#71C5E8",
  inverse_header_color = "#FFFFFF",
  background_color = "#EAE9EA",
  link_color = "#71C5E8",
  # easy to fetch colors
  colors = c( 
    white = "#FFFFFF",
    green = "#336666",
    lblue = "#71C5E8"
    )
)
```

```{r other-options}
library(tidyverse)
library(kableExtra)
library(fontawesome)

# ggplot global options
theme_set(theme_bw(base_size = 20))

# Democracy scores maps
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(countrycode)
library(gridExtra)
```

class: inverse

## Reminder

- Quiz 1 opens on Thursday, October 7 at noon
- Due on Friday, October 8 at 5:00 PM
- Weeks 1-4

---

## Last time

- We study politics within countries, by which we usually mean nation-states
- Two explanations for the emergence of modern nation-states:

    1. Contractarian view
    2. Predatory view
    
- They also help us understand variation in **state capacity**, **state fragility**, or **stateness** across countries and over time

- This is the foundation for many other interesting questions

---

## The way forward

- Why are some countries `democracies` and other `dictatorships`?

- Which `regime` produces better economic performance?

- What factors influence democratic survival?

&nbsp;

- To answer these questions, we need to classify countries as **democracies** or **dictatorships**

---
class: inverse

## Aside: English is hard

- **State:** An organization with comparative advantage in violence

- **Government:** The group of people who rule the state

- **Regime:** Institutions that determine how governments are formed

--

&nbsp;

.center[This week is about **regime types**]

---
layout:true 
## Democracy is an old concept

---
- ... and it was a bad thing!

- **Demokratia:** Rule by the `(common)` people

- **Plato:** Government by the poor, uneducated against the rich, educated

- Would lead to mob rule and class warfare

---

- **Aristotle:** All forms of government are corruptible

```{r}
aristotle = data.frame(
  number = c("One", "Few", "Many"),
  good = c("Monarchy", "Aristocracy", "Politeia"),
  bad = c("Tyranny", "Olygarchy", "Democracy")
)

colnames(aristotle) = c("Number of rulers", "Good for all", "Good for the rulers")

aristotle %>% 
  kbl() %>% 
  add_header_above(c(" " = 1, "Good form" = 1, "Bad form" = 1))
```

- **Democracy:** Rule by the worst class (Monarchy preferred)
- **Until 18th Century:** 

    - No elections `(aristocracy)`, but `lottery`
    - No representation, but `direct legislation`
    
- `Democracy` $=$ `representative government` only after French and American Revolutions

---

- As demands for representative government increased, the classical 3-way distinction disappeared in favor of a 2-way distinction between **democracy** and **dictatorship**

- This is also when democracy started to associate with `equality`, `justice`, and `freedom`

---
layout:false

## Defining democracy

- These are **substantive** definitions of democracy (and dictatorship) since they focus on what it `should accomplish` rather than what `it is`

--

- **Robert Dahl** proposed a **minimalist** definition instead

- He used the term **polyarchy** to distinguish it from conventional use. Also to emphasize that this as an ideal type, and that different regimes can get closer or further from the ideal

---

## Polyarchy

- Regimes can be categorized in two dimensions:

    1. **Contestation:** Extent to which citizens can organize into competing blocks to press for policy change
    2. **Inclusion:** Who gets to participate

```{r}
polyarchy = data.frame(
  countries = c("Apartheid\nSouth Africa", "China", "Soviet Union", "United States", "Polyarchy\n(Ideal type)"),
  Inclusion = c(10, 30, 60, 70, 90),
  Contestation = c(70, 10, 10, 70, 90)
)

ggplot(polyarchy) +
  aes(x = Inclusion, y = Contestation, label = countries) +
  geom_text(size = 5) +
  scale_x_continuous(breaks = c(10, 90), labels = c("Low", "High")) +
  scale_y_continuous(breaks = c(10, 90), labels = c("Low", "High")) +
  coord_cartesian(xlim = c(-5, 105), ylim = c(-5, 105)) +
  theme_xaringan() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
 

```


---

## Polyarchy

- This definition is **minimalist** because its sole purpose is to classify regimes

- We want this because later on we want to learn about the causes and consequences of regime types

- `Contestation` and `inclusion` are abstract concepts

- How do we turn them into a concrete measure of democracy?

---

## From concept to measurement

- **Concept:** Mental category that conveys meaning. Like `theories` or `models` they are always wrong, but sometimes useful

--

- **Measure/indicator:** An observable representation of a concept

    - CGG say a `quantification`, but a measure need not be a "quantitiy" `r fa("fire-alt")`
    - `Example:` Hot and cold are categorial measures of temperature
    
--

- **Operationalization:** The process by which we translate a concept into a measure. It often implies justifying why the measure is a good representation of the concept

---

## Measures of Democracy

.pull-left[

### Substantive
- Freedom House [`r fa("external-link-alt")`](https://freedomhouse.org/)

- Varieties of Democracy [`r fa("external-link-alt")`](https://www.v-dem.net/en/)
]

.pull-right[

### Minimalist
- Polity IV-V [`r fa("external-link-alt")`](http://www.systemicpeace.org/polityproject.html)

- Democracy-Dictatorship Index [`r fa("external-link-alt")`](https://sites.google.com/site/joseantoniocheibub/datasets/democracy-and-dictatorship-revisited)
]

--

&nbsp;

.center[Let's talk about these in logical order]

---
## Freedom House

- Based on two dimensions:

    1. Political rights (10 questions, scored 0-4)
    2. Civil rights (15 questions, scored 0-4)

- Both transformed to a 7 point scale
- A country's FH is the average of political and civil rights
- A `substantive` measure because democracy should promote freedom

---

## Polity IV-V

- $Polity = Democracy - Autocracy$

- Ranges from -10 to 10

- Five dimensions:

    1. Competitiveness of executive recruitment
    2. Openness of executive recruitment
    3. Regulation of political participation
    4. Competitiveness of political participation
    5. Executive constraints

- `Minimalist` since 1 to 4 capture `inclusion` and `contestation`
- 5 adds the idea that governments must be limited


---

## Democracy-Dictatorship Index (DD)

- Doubles down on the `minimalist` approach

- Claims the core feature of democracy is `contestation`

- A country is a `democracy` if:

    1. The chief executive is elected
    2. The legislature is elected
    3. There is more than one party competing in elections
    4. An alternation in power under identical electoral rules has taken place
    
- If any of these conditions does not hold, the country is a `dictatorship`

- This is a **dichotomous* measure



---

## Varieties of Democracy (V-DEM)

HERE

---
## What's a good measure?

- Reliable, valid


---
layout: true

## Does it matter?

---

```{r}
load("data/dem_scores.RData")

# fix polity score
polity = polity %>% mutate(polity = ifelse(abs(polity) > 10, NA, polity))

# Merge with map
world = ne_countries(scale = "small", returnclass = "sf")

fh_map = left_join(world, fh, by = c("iso_a3" = "iso3c"))
poli_map = left_join(world, polity, by = c("iso_a3" = "iso3c"))
dd_map = left_join(world, dd, by = c("iso_a3" = "iso3c"))
vdem_map = left_join(world, vdem, by = c("iso_a3" = "iso3c"))

fhp = ggplot(fh_map) +
  geom_sf(aes(fill = fh_total_reversed)) +
  labs(fill = "") +
  theme_void() +
  theme(legend.position = "bottom") +
  scale_fill_viridis_c()

polip = ggplot(poli_map) +
  geom_sf(aes(fill = polity)) +
  theme_xaringan() +
  labs(fill = "") +
  theme_void() +
  theme(legend.position = "bottom") + 
  scale_fill_viridis_c()

ddp = ggplot(dd_map) +
  geom_sf(aes(fill = as.factor(dd))) +
  theme_xaringan() +
  labs(fill = "") +
  theme_void() +
  theme(legend.position = "bottom") +
  scale_fill_viridis_d(labels = c("Dictatorship", "Democracy", "NA"))

vdemp = ggplot(vdem_map) +
  geom_sf(aes(fill = v2x_polyarchy)) +
  theme_xaringan() +
  labs(fill = "") +
  theme_void() +
  theme(legend.position = "bottom") +
  scale_fill_viridis_c()
```

### Freddom House 2020 (reversed)

```{r, fig.dim = c(6,3)}
fhp
```

---

### Polity V 2018

```{r, fig.dim = c(6,3)}
polip
```

---

### DD Index 2008

```{r, fig.dim = c(6,3)}
ddp
```

---

### V-DEM Polyarchy score 2020

```{r, fig.dim = c(6,3)}
vdemp
```

---

### Comparing scores

HERE

---
layout: false

## Takeaways

HERE


---
class: inverse center middle

# Reminder:
## Quiz 1 Due Friday by 5:00 PM

# Next Week:
## Economic Determinants of Democracy and Dictatorship