---
title: "Economic Determinants of Democracy and Dictatorship"
subtitle: "Comparative Politics"  
date: "October 11-15, 2021"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%/%total%" 
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)

library(knitr)
opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  title_slide_background_image = "figs/logo.png",
  title_slide_background_size = "8%",
  title_slide_background_position = "50% 95%",
  primary_color = "#336666",
  secondary_color = "#71C5E8",
  inverse_header_color = "#FFFFFF",
  background_color = "#EAE9EA",
  link_color = "#71C5E8",
  # easy to fetch colors
  colors = c( 
    white = "#FFFFFF",
    green = "#336666",
    lblue = "#71C5E8"
    )
)
```

```{r other-options}
library(tidyverse)
library(kableExtra)
library(fontawesome)

# ggplot global options
theme_set(theme_bw(base_size = 20))

# Plot democracy and WDI data
library(countrycode)
library(WDI)

```

## Last time

- We discussed classifying regime types

    - Continuous or categorical
    - Minimalist or substantive

- This all matters because we want to understand the consequences of being a democracy or a dictatorship

- We usually promote democracy because we believe it brings good outcomes

- But how do countries become democracies?

- And once they do, how do they stay democracies?

- This week is about the **economic** explanations for regime change and survival

- Next week is about the **cultural** explanations

---

## Modernization theory

- All societies pass through the same stages of economic development

.pull-left[
### Traditional society

- Large agriculture
- Small industry
- Small service
- Dictatorship
]

.pull-right[
### Modern society

- Small agriculture
- Large industry
- Large service
- Democracy
]

- As societies develop, they are `more likely` to become a democraty
- So we say that *development* causes *democracy*

---
class: inverse 

## Causal statements in the social sciences

.center[
### Economic development $\rightarrow$ Democracy
]

--

.center[
### Explanatory variable $\rightarrow$ Outcome variable
]

--

.center[
### X $\rightarrow$ Y
]

--

- We do not see `causation` happening, only `association`

--

- So just because richer countries tend to be democracies it does not mean that development `causes` democracy

--

- It could be the other way around

--

- Or a big coincidence!

--

- Hence the importance of finding **critical tests** for our theories

---

## Using DD index

```{r}
load("data/dem_scores.RData")

gdp = WDI(indicator = "NY.GDP.PCAP.KD", country = "all", start = 2007, end = 2020)

# some recoding
gdp$iso3c = countrycode(gdp$iso2c, origin = "iso2c", destination = "iso3c")

gdp$gdp_pc = gdp$NY.GDP.PCAP.KD/1000

# merge
dd_gdp = left_join(dd, gdp %>% filter(year == 2007), by = "iso3c")

# visualize
# DO DENSITIES DIVIDED ON WHETHER WE HAVE DEMOCRACY OR DICTATORSHIP
```



