---
title: "Democracy and Dictatorship"
subtitle: "Comparative Politics"  
date: "October 4-8, 2021"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%/%total%" 
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)

library(knitr)
opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  title_slide_background_image = "figs/logo.png",
  title_slide_background_size = "8%",
  title_slide_background_position = "50% 95%",
  primary_color = "#336666",
  secondary_color = "#71C5E8",
  inverse_header_color = "#FFFFFF",
  background_color = "#EAE9EA",
  link_color = "#71C5E8",
  # easy to fetch colors
  colors = c( 
    white = "#FFFFFF",
    green = "#336666",
    lblue = "#71C5E8"
    )
)
```

```{r other-options}
library(tidyverse)
library(kableExtra)
library(fontawesome)

# ggplot global options
theme_set(theme_bw(base_size = 20))
```

class: inverse

## Reminder

- Quiz 1 opens on Thursday, October 7 at noon
- Due on Friday, October 8 at 5:00 PM
- Weeks 1-4

---

## Last time

- We study politics within countries, by which we usually mean nation-states
- Two explanations for the emergence of modern nation-states:

    1. Contractarian view
    2. Predatory view
    
- They also help us understand variation in **state capacity**, **state fragility**, or **stateness** across countries and over time

- This is the foundation for many other interesting questions

---

## The way forward

- Why are some countries `democracies` and other `dictatorships`?

- Which `regime` produces better economic performance?

- What factors influence democratic survival?

&nbsp;

- To answer these questions, we need to classify countries as **democracies** or **dictatorships**

---
layout:true 
## Democracy is an old concept

---
- ... and it was a bad thing!

- **Demokratia:** Rule by the `(common)` people

- **Plato:** Government by the poor, uneducated against the rich, educated

- Would lead to mob rule and class warfare

---

- **Aristotle:** All forms of government are corruptible

```{r}
aristotle = data.frame(
  number = c("One", "Few", "Many"),
  good = c("Monarchy", "Aristocracy", "Politeia"),
  bad = c("Tyranny", "Olygarchy", "Democracy")
)

colnames(aristotle) = c("Number of rules", "Good for all", "Good for the rulers")

aristotle %>% 
  kbl() %>% 
  add_header_above(c(" " = 1, "Good form" = 1, "Bad form" = 1))
```

- **Democracy:** Rule by the worst class (Monarchy preferred)
- **Until 18th Century:** 

    - No elections `(aristocracy)`, but `lottery`
    - No representation, but `direct legislation`
    
- `Democracy` $=$ `representative government` only after French and American Revolutions

---

- As demands for representative government increased, the classical 3-way distinction disappeared in favor of a 2-way distinction between **democracy** and **dictatorship**

- This is also when democracy started to associate with `equality`, `justice`, and `freedom`

---
layout:false

## Defining democracy

[Substantive vs. minimalist]

## From concept to measurement

[Clarify social science and then illustrate. Then measures of democracy.]

