% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  11pt,
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
  \setmainfont[]{cochineal}
  \setmonofont[]{Fira Code}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}
    }
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\author{Gustavo Diaz}
\date{January 05, 2022}

% Jesus, okay, everything above this comment is default Pandoc LaTeX template. -----
% ----------------------------------------------------------------------------------
% I think I had assumed beamer and LaTex were somehow different templates.


\usepackage{kantlipsum}

\usepackage{abstract}
\renewcommand{\abstractname}{}    % clear the title
\renewcommand{\absnamepos}{empty} % originally center

\renewenvironment{abstract}
 {{%
    \setlength{\leftmargin}{0mm}
    \setlength{\rightmargin}{\leftmargin}%
  }%
  \relax}
 {\endlist}

\makeatletter
\def\@maketitle{%
  \newpage
%  \null
%  \vskip 2em%
%  \begin{center}%
  \let \footnote \thanks
      {\fontsize{18}{20}\selectfont\raggedright  \setlength{\parindent}{0pt} \@title \par}
    }
%\fi
\makeatother


 



%\author{\Large \vspace{0.05in} \newline\normalsize\emph{}  }


\date{}

\usepackage{titlesec}

% 
\titleformat*{\section}{\large\bfseries}
\titleformat*{\subsection}{\normalsize\itshape} % \small\uppercase
\titleformat*{\subsubsection}{\normalsize\itshape}
\titleformat*{\paragraph}{\normalsize\itshape}
\titleformat*{\subparagraph}{\normalsize\itshape}

% add some other packages ----------

% \usepackage{multicol}
% This should regulate where figures float
% See: https://tex.stackexchange.com/questions/2275/keeping-tables-figures-close-to-where-they-are-mentioned
\usepackage[section]{placeins}



\makeatletter
\@ifpackageloaded{hyperref}{}{%
\ifxetex
  \PassOptionsToPackage{hyphens}{url}\usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \PassOptionsToPackage{hyphens}{url}\usepackage[draft,unicode=true]{hyperref}
\fi
}

\@ifpackageloaded{color}{
    \PassOptionsToPackage{usenames,dvipsnames}{color}
}{%
    \usepackage[usenames,dvipsnames]{color}
}
\makeatother
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={ ()},
             pdfkeywords = {},  
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

% Add an option for endnotes. -----



% This will better treat References as a section when using natbib
% https://tex.stackexchange.com/questions/49962/bibliography-title-fontsize-problem-with-bibtex-and-the-natbib-package



% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother



\pagenumbering{gobble}

\newtheorem{hypothesis}{Hypothesis}

\usepackage{fontawesome}

\newcommand{\blankline}{\quad\pagebreak[2]}
\usepackage{graphicx}

\begin{document}



\hfill
\begin{minipage}[t]{1\textwidth}
\raggedleft%
{\bfseries Gustavo Diaz }\\[.35ex]
\emph{\small Center for Inter-American Policy and Research\\
Tulane University} \\[.35ex]
\faPhone \hspace{1 mm} \small{+1 217 904 0581} \\ 
\faEnvelopeO \hspace{1 mm} \small{\tt \href{mailto:gustavodiaz@tulane.edu}{\nolinkurl{gustavodiaz@tulane.edu}}} \\ 
\faGlobe \hspace{1 mm} \small{\href{http://gustavodiaz.org}{\tt gustavodiaz.org}}\\ 
\hspace{1cm} \\
 January 05, 2022 \\ 
\end{minipage}

% \vspace*{1em} 

\vspace*{1em}

Dear Members of the Search Committee,

\vspace*{1em}

 

% January 05, 2022
% 
% \vspace*{1em} 
% 

% % \setlength{\parindent}{16pt}
% \setlength{\parskip}{0pt}
% 
I write to express my interest in your call for an Assistant Professor
in Methodology and Statistics. I am a postdoctoral fellow at the Center
for Inter-American Policy and Research at Tulane University. I received
my PhD from the Department of Political Science at the University of
Illinois at Urbana-Champaign. I research the application of tools from
machine learning and design-based causal inference to improve research
design in experiments and observational studies. I base my insights on
my substantive work on accountability and representation in the Global
South.

My primary agenda focuses on improving our tools for understanding
hard-to-observe social and political phenomena along two research
programs. The first research program focuses on analyzing interference
effects in settings where the researcher observes units' placement in a
network, but not how they connect to each other. I propose a model
selection protocol that overcomes this difficulty. The key is to use a
distance metric to identify alternative model specifications that follow
from a theory, and then choose among them using a model selection
technique of choice. I illustrate the usefulness of this approach in the
context of supervised learning model selection by revisiting the
findings of an experiment on the spillover effects of election observers
on voter registration irregularities in Ghana.

The second research program focuses on navigating the bias-variance
trade-off in survey and field experiments. In a chapter with Christopher
Grady (USAID) and Jim Kuklinski (Illinois) in \emph{The SAGE Handbook of
Research Methods in Political Science and International Relations}, we
discuss the merits and challenges of increased complexity in the design
of survey experiments. I bring the lessons of our chapter to a
solo-authored piece under review that introduces new tools to assess the
validity of findings in double list experiments. This variant of the
list experiment promises more precise prevalence estimates but comes
with under-explored questionnaire design complications. In a working
paper with Jake Bowers (Illinois) and Christopher Grady, we discuss the
circumstances under which researchers should prefer biased yet precise
estimators to analyze experimental data, including applications to
block-randomization, covariance adjustment, and M-estimation.

In my time at Illinois, I served as a teaching assistant for several
statistics courses at the undergraduate and PhD levels using a flipped
classroom approach. I also served as a math camp instructor for incoming
graduate students for three consecutive years and started a
collaborative project in which graduate students introduced their peers
to new methodological tools. I am prepared to teach courses on causal
inference, computational social science, and machine learning. You can
find copies of current and sample syllabi in my website.

I believe my expertise would make me an excellent fit at Utrecht
University. Please find enclosed the materials supporting my
application. If you have any questions, you can contact me via e-mail or
phone.

\end{document}
